SNIK: Textbook Knowledge on Hospital Information Management as Linked Data

### 2021-11-21 paper.tex
Nearly finished alpha.

% add. options: [seceqn,secthm,crcready,onecolumn]
%\documentclass[sw]{iosart2x}
\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
%\usepackage{dcolumn}
%\usepackage{endnotes}
\usepackage{booktabs}
\usepackage{tabulary}
\usepackage{tabularx}
\usepackage{csquotes}
\usepackage{siunitx}
\usepackage{todonotes}
%\usepackage{natbib} % does not seem to work with ios 1
\newcommand{\citet}{\cite}% citet is not defined without natbib
\newcommand{\citep}{\cite}% citep is not defined without natbib
\setlength{\marginparwidth}{2.1cm}% enough space for todonotes
\usepackage{listings}
\usepackage{aurl}
\daurl{ob}{http://www.snik.eu/ontology/bb/}
\daurl{bb}{http://www.snik.eu/ontology/bb/}
\daurl{meta}{http://www.snik.eu/ontology/meta/}
\lstset{language=SPARQL,breaklines=true}
\usepackage[noabbrev,capitalize]{cleveref}
\newcommand{\snikversion}{1.3.0}
\newcommand{\sniktriples}{72046}
\newcommand{\snikclasses}{2655}
\newcommand{\snikproperties}{339}
\newcommand{\sniklinks}{581}
%%%%%%%%%%% End of definitions

@@

\begin{table}
\caption{Statistics}
\label{tab:statistics}
\centering
\begin{tabulary}{\columnwidth}{lS}
\toprule
\textbf{Category}		&\textbf{Count}\\
\midrule
Triples				&\sniktriples{}\\
Classes				&\snikclasses{}\\
Properties			&\snikproperties{}\\
Internal Links		&360\\
Links to HITO  		&260\\
Links to DBpedia 	&\sniklinks{}\\
%Subontologies	&\url{http://www.snik.eu/ontology/bb}\\
%		&\url{http://www.snik.eu/ontology/ciox,}\\
%		&\url{http://www.snik.eu/ontology/ob}\\
%		&\url{http://www.snik.eu/ontology/he}\\
%		&\url{http://www.snik.eu/ontology/it4it}\\
\bottomrule
\end{tabulary}%The source code for the services is available at \url{https://github.com/imise}
\end{table}

@@

\section{Data Transformation Process}
\subsection{Interlinking}
%As shown in \cref{...}, we aligned our dataset to several others:
We manually aligned our dataset to HITO, the Health IT Ontology~\citep{hitometh2}
We automatically aligned our dataset to:
- DBpedia~\citep{dbpedia}, as  
- 
\section{Lifecycle}\label{sec:application}

@@

\paragraph{Compound Layout}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{img/combine.pdf}
    \caption{Grouping of equivalent concepts from different sources in SNIK Graph (color coded).}% using the \enquote{combine matches} option.}
	\label{fig:combine}
\end{figure}
\vspace{-3pt}
%\noindent For citations of references, we prefer the use of square brackets and consecutive numbers, e.g. as shown by Author et al. [2], [3, pp. 5–10], as mentioned earlier [1], [3], [9]. The following bibliography provides the basic formats as a reference list with entries for journal articles [1], book chapter [2], as well as a URL [5]. For further guidance please refer to the \href{https://ieeeauthorcenter.ieee.org/wp-content/uploads/IEEE-Reference-Guide.pdf}{IEEE-Reference-Guide}.

@@

\subsection{Neighbourhood Operations}
Exploration using neighbours, that is the successive uncovering of nodes adjacent to a starting node given by a user, is a common feature of tools such as \cite{lodlive} and \cite{vizlod}.%
\subsection{Star}
The directed and undirected star operations show nodes connected to selected nodes via all paths that contain at most one property other than \aurl{skos}{closeMatch}.
The \emph{circle star} also rearranges the nodes using the force-directed layout locally on the currently visible subgraph.
\Cref{fig:star} shows a mind map of strategic information management, created by an undirected star, which can be used by a teacher to prepare a lecture about the topic.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\linewidth]{img/strategic-im.pdf}
    \caption{Undirected \emph{Star} of \aurl{bb}{StrategicInformationManagement}.}
	\label{fig:star}
\end{figure}
\vspace{-3pt}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\linewidth]{img/spiderworm.pdf}
    \caption{\emph{Spiderworm} from \aurl{bb}{PatientIdentificationNumber} to \aurl{bb}{QualityOfData}~\cite{snikgraphposter}.}
	\label{fig:spiderworm}
\end{figure}
\vspace{-3pt}

@@

\begin{figure}
\caption{Students learn new concepts about HIS quality by linking them to concepts already learned.
A teacher asks a student to find out how the new concept \enquote{Quality of Data} is linked to the \enquote{Patient Identification Number}.
The student connects the two concepts by using the “spiderworm“ visualization and learns that a patient identification number is associated with object identity.
Object identity is a subclass of integrity of data.
Besides integrity of data, there are also 13 other criteria for quality of data.}
\label{fig:snik-graph-spiderworm}
\includegraphics[width=\columnwidth]{img/snik-graph-spiderworm.png}
\end{figure}
\todo{FJ: In diesem Abschnitt fehlt die Diskussion, wie es durch die verschiedenen Zielgruppen bei Use Cases genutzt werden kann.}
SNIK Graph is a web application\footnote{\url{http://www.snik.eu/graph}} that transforms the classes and triples of SNIK to nodes and edges of a graph.
The graph is visualized using the Cytoscape.js~\citep{cytoscape} library with the force-directed Euler layout, see~\cref{fig:snik-graph-overview}.
With several thousand classes, specific parts of SNIK can be hard to discern.
Thus, there are several options to view subgraphs of SNIK, for example to show only a specific chapter of a book to prepare a lecture about a specific topic.
Users can also search for and restrict the view to a single class and then show the neighbourhood of that class (see \cref{fig:snik-graph-circle-star}) and subsequently the neighbourhood of selected neighbours of the previous step.
SNIK Graph can also calculate the shortest path to between classes.
We also tried to automatically find the most interesting path but this was not successfull as it is subjective and the depends on the goal of the user.
Path and neighourhood operations are joined in the \enquote{spider worm} (see \cref{fig:snik-graph-spiderworm}), which consists of the shortest path between a start node and an end node together with the end node’s neighbourhood, illustrate the context of a concept.
\fi
\subsection{Quiz}
The DBpedia-powered Clover Quiz~\citep{cloverquiz} shows that ontologies can be used to automatically generate multiple-choice questions.
We apply this approach to SNIK and generate 1231 English questions using the templates described in \cref{tab:templates}.
It was used by students from the Universities of Amsterdam, Heidelberg and Leipzig during the international Frank van Swieten lectures in 2019.
To provide difficult wrong answers, the so called \emph{distractors}, we use direct neighbours of the classes that represent the correct answers.
Due to the limited nesting capabilties of SPARQL, which does not support loops but only subqueries that cannot access variables declared outside their scope, querying for distinct sets of 4 neighbours, where exactly one was connected over a certain property to a certain object, was not possible with our Virtuoso SPARQL endpoint initially.
We circumvented the timeout by calculating the neighbour-relation as a first step and uploading it to the endpoint.

@@

\begin{table*}
\caption{Templates}
\label{tab:templates}
\centering
\begin{tabulary}{\textwidth}{lL}
\toprule
\textbf{Template}	&\textbf{Description}\\
\midrule
Definition		&Ask for the class that fits the given textbook definition.\\
Example			&What is defined as \enquote{Examination of in and out patients in radiological department}?\\
Distractors		&Labels of other classes (that have a path of length 2 or less to the correct class.)\\
\midrule
Subject			&Ask for the class that is related via a given relation to a given object.\\
Example			&Who is \emph{involved in} a \emph{healthcare network}?\\
Distractors		&Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object.\\
\bottomrule
\end{tabulary}
\end{table*}

%The technical environment of SNIK is characterized in previous work [11].
\section{Discussion}
SNIK is based on the meta model from Figure \todo{reference}.
Thus, SNIK can be considered as an archetype for ontologies describing for a given domain (here \enquote{HIS management}), what functions a certain role has to carry out and what information a person of this role needs and what information she/he provides while carrying out the functions.
This may be used for domains like the medical department of a hospital as well.\todo{FJ: Ich glaube, das müsste man ausführlicher erklären, das würde ich eher ans Ende der Diskussion nehmen.}
This would be valuable knowledge, for instance as basis for systems analysis projects [1].
Different user groups can use the interfaces of SNIK that are most suitable to them.
For example, the IT management of a hospital can use SNIK as a vocabulary to integrate data from different formats that result from application components from different vendors.
Using the SPARQL1 query language, experts can also integrate applications with the SNIK ontology.
 Experts can also use the SPARQL query editor to answer specific questions, see Table 2.
Students and teachers, on the other hand, benefit more from SNIK Graph, which intuitively presents knowledge without the need for technical skills.

Figure 4– A \enquote{Spider Worm} between the classes \enquote{Change Management} and \enquote{HIS Quality} in the SNIK Graph visualization web application.
The content of SNIK is taken from textbooks which are protected by property rights of publishing houses or taken from an interview with a CIO holding property rights as well.
Making the SNIK ontology openly available may therefore conflict with these property rights.
In case of SNIK, we appreciate the approvals of the publishing houses Springer, de Gruyter and Schattauer and of the CIO.
In general, missing approvals may prevent the creation of such open ontologies and thus the provision of textbook content as open knowledge.
The approvals for SNIK were possible because for each class and relation being part of SNIK there is a clear indication of its source (see code \enquote{bb} and chapter information in Table 1).
In future work we will use these indications for establishing hyperlinks into the sources such that users of SNIK may use SNIK as a smart index for retrieving more detailed knowledge and supporting illustrations from the original textbooks – if digitally available.
We are convinced that this way an ontology like SNIK is not a competitor of textbooks but a smart guide for using a textbook.
Thus, it might be attractive for publishing houses.

\todo{verweis auf existierende papers}
\todo{benutzung durch braunschweig}
\todo{benutzung durch amsterdam}

>>>>

% add. options: [seceqn,secthm,crcready,onecolumn]
%\documentclass[sw]{iosart2x}
\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
%\usepackage{dcolumn}
%\usepackage{endnotes}
\usepackage{booktabs}
\usepackage{tabulary}
\usepackage{tabularx}
\usepackage{csquotes}
\usepackage{siunitx}
\usepackage[disable]{todonotes}
%\usepackage{natbib} % does not seem to work with ios 1
\newcommand{\citet}{\cite}% citet is not defined without natbib
\newcommand{\citep}{\cite}% citep is not defined without natbib
\setlength{\marginparwidth}{2.1cm}% enough space for todonotes
\usepackage{listings}
\usepackage{aurl}
\daurl{ob}{http://www.snik.eu/ontology/bb/}
\daurl{bb}{http://www.snik.eu/ontology/bb/}
\daurl{meta}{http://www.snik.eu/ontology/meta/}
\lstset{language=SPARQL,breaklines=true}
\usepackage[noabbrev,capitalize]{cleveref}
\newcommand{\snikversion}{1.3.0}
\newcommand{\sniktriples}{72046}
\newcommand{\snikclasses}{2655}
\newcommand{\snikproperties}{339}
\newcommand{\sniklinks}{581}
%%%%%%%%%%% End of definitions

@@

\begin{table}
\caption{Statistics}
\label{tab:statistics}
\centering
\begin{tabulary}{\columnwidth}{lS}
\toprule
\textbf{Category}					&\textbf{Count}\\
\midrule
Triples								&\sniktriples{}\\
Classes								&\snikclasses{}\\
Properties							&\snikproperties{}\\
Internal Links						&360\\
Links to HITO  						&260\\
Links to DBpedia~\citep{dbpedia} 	&\sniklinks{}\\
%Subontologies	&\url{http://www.snik.eu/ontology/bb}\\
%		&\url{http://www.snik.eu/ontology/ciox,}\\
%		&\url{http://www.snik.eu/ontology/ob}\\
%		&\url{http://www.snik.eu/ontology/he}\\
%		&\url{http://www.snik.eu/ontology/it4it}\\
\bottomrule
\end{tabulary}%The source code for the services is available at \url{https://github.com/imise}
\end{table}

@@

%\section{Data Transformation Process}
%\subsection{Interlinking}
%As shown in \cref{...}, we aligned our dataset to several others:
%We manually aligned our dataset to HITO, the Health IT Ontology~\citep{hitometh2}
%We automatically aligned our dataset to:
%- DBpedia~\citep{dbpedia}, as  
%- 
\section{Lifecycle}\label{sec:application}

@@

\paragraph{Compound Layout}
\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{img/combine.pdf}
    \caption{Grouping of equivalent concepts from different sources in SNIK Graph (color coded).}% using the \enquote{combine matches} option.}
	\label{fig:combine}
\end{figure}
\vspace{-3pt}
%\noindent For citations of references, we prefer the use of square brackets and consecutive numbers, e.g. as shown by Author et al. [2], [3, pp. 5–10], as mentioned earlier [1], [3], [9]. The following bibliography provides the basic formats as a reference list with entries for journal articles [1], book chapter [2], as well as a URL [5]. For further guidance please refer to the \href{https://ieeeauthorcenter.ieee.org/wp-content/uploads/IEEE-Reference-Guide.pdf}{IEEE-Reference-Guide}.

@@

\subsection{Neighbourhood Operations}
Exploration using neighbours, that is the successive uncovering of nodes adjacent to a starting node given by a user, is a common feature of tools such as \cite{lodlive} and \cite{vizlod}.%
\subsection{Star}
The directed and undirected star operations show nodes connected to selected nodes via all paths that contain at most one property other than \aurl{skos}{closeMatch}.
The \emph{circle star} also rearranges the nodes using the force-directed layout locally on the currently visible subgraph.
\Cref{fig:star} shows a mind map of strategic information management, created by an undirected star, which can be used by a teacher to prepare a lecture about the topic.
\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{img/strategic-im.pdf}
    \caption{Undirected \emph{Star} of \aurl{bb}{StrategicInformationManagement}.}
	\label{fig:star}
\end{figure}
\vspace{-3pt}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{img/spiderworm.pdf}
    \caption{\emph{Spiderworm} from \aurl{bb}{PatientIdentificationNumber} to \aurl{bb}{QualityOfData}~\cite{snikgraphposter}.}
	\label{fig:spiderworm}
\end{figure}
\vspace{-3pt}

@@

\begin{figure}
\caption{Students learn new concepts about HIS quality by linking them to concepts already learned.
A teacher asks a student to find out how the new concept \enquote{Quality of Data} is linked to the \enquote{Patient Identification Number}.
The student connects the two concepts by using the “spiderworm“ visualization and learns that a patient identification number is associated with object identity.
Object identity is a subclass of integrity of data.
Besides integrity of data, there are also 13 other criteria for quality of data.}
\label{fig:snik-graph-spiderworm}
\includegraphics[width=\columnwidth]{img/snik-graph-spiderworm.png}
\end{figure}
\todo{FJ: In diesem Abschnitt fehlt die Diskussion, wie es durch die verschiedenen Zielgruppen bei Use Cases genutzt werden kann.}
SNIK Graph is a web application\footnote{\url{http://www.snik.eu/graph}} that transforms the classes and triples of SNIK to nodes and edges of a graph.
The graph is visualized using the Cytoscape.js~\citep{cytoscape} library with the force-directed Euler layout, see~\cref{fig:snik-graph-overview}.
With several thousand classes, specific parts of SNIK can be hard to discern.
Thus, there are several options to view subgraphs of SNIK, for example to show only a specific chapter of a book to prepare a lecture about a specific topic.
Users can also search for and restrict the view to a single class and then show the neighbourhood of that class (see \cref{fig:snik-graph-circle-star}) and subsequently the neighbourhood of selected neighbours of the previous step.
SNIK Graph can also calculate the shortest path to between classes.
We also tried to automatically find the most interesting path but this was not successfull as it is subjective and the depends on the goal of the user.
Path and neighourhood operations are joined in the \enquote{spider worm} (see \cref{fig:snik-graph-spiderworm}), which consists of the shortest path between a start node and an end node together with the end node’s neighbourhood, illustrate the context of a concept.
\fi
\subsection{Quiz}

\begin{table}[t]
\caption{Initial Quiz question templates.}
\label{tab:templates}
\centering
\begin{tabulary}{\columnwidth}{lL}
\toprule
\textbf{Template}	&\textbf{Description}\\
\midrule
Definition		&Ask for the class that fits the given textbook definition.\\
Example			&What is defined as \enquote{Examination of in and out patients in radiological department}?\\
Distractors		&Labels of other classes (that have a path of length 2 or less to the correct class.)\\
\midrule
Subject			&Ask for the class that is related via a given relation to a given object.\\
Example			&Who is \emph{involved in} a \emph{healthcare network}?\\
Distractors		&Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object.\\
\bottomrule
\end{tabulary}
\end{table}

The DBpedia-powered Clover Quiz~\citep{cloverquiz} shows that ontologies can be used to automatically generate multiple-choice questions.
We apply this approach to SNIK and generate 1231 English questions using the templates described in \cref{tab:templates}.
It was used by students from the Universities of Amsterdam, Heidelberg and Leipzig during the international Frank van Swieten lectures in 2019.
To provide difficult wrong answers, the so called \emph{distractors}, we use direct neighbours of the classes that represent the correct answers.
Due to the limited nesting capabilties of SPARQL, which does not support loops but only subqueries that cannot access variables declared outside their scope, querying for distinct sets of 4 neighbours, where exactly one was connected over a certain property to a certain object, was not possible with our Virtuoso SPARQL endpoint initially.
We circumvented the timeout by calculating the neighbour-relation as a first step and uploading it to the endpoint.

@@

%The technical environment of SNIK is characterized in previous work [11].
\iffalse
\section{Discussion}
SNIK is based on the meta model from Figure \todo{reference}.
Thus, SNIK can be considered as an archetype for ontologies describing for a given domain (here \enquote{HIS management}), what functions a certain role has to carry out and what information a person of this role needs and what information she/he provides while carrying out the functions.
This may be used for domains like the medical department of a hospital as well.\todo{FJ: Ich glaube, das müsste man ausführlicher erklären, das würde ich eher ans Ende der Diskussion nehmen.}
This would be valuable knowledge, for instance as basis for systems analysis projects [1].
Different user groups can use the interfaces of SNIK that are most suitable to them.
For example, the IT management of a hospital can use SNIK as a vocabulary to integrate data from different formats that result from application components from different vendors.
Using the SPARQL query language, experts can also integrate applications with the SNIK ontology.
 Experts can also use the SPARQL query editor to answer specific questions, see Table 2.
Students and teachers, on the other hand, benefit more from SNIK Graph, which intuitively presents knowledge without the need for technical skills.

Figure 4– A \enquote{Spider Worm} between the classes \enquote{Change Management} and \enquote{HIS Quality} in the SNIK Graph visualization web application.
The content of SNIK is taken from textbooks which are protected by property rights of publishing houses or taken from an interview with a CIO holding property rights as well.
Making the SNIK ontology openly available may therefore conflict with these property rights.
In case of SNIK, we appreciate the approvals of the publishing houses Springer, de Gruyter and Schattauer and of the CIO.
In general, missing approvals may prevent the creation of such open ontologies and thus the provision of textbook content as open knowledge.
The approvals for SNIK were possible because for each class and relation being part of SNIK there is a clear indication of its source (see code \enquote{bb} and chapter information in Table 1).
In future work we will use these indications for establishing hyperlinks into the sources such that users of SNIK may use SNIK as a smart index for retrieving more detailed knowledge and supporting illustrations from the original textbooks – if digitally available.
We are convinced that this way an ontology like SNIK is not a competitor of textbooks but a smart guide for using a textbook.
Thus, it might be attractive for publishing houses.
\fi
\todo{verweis auf existierende papers}
\todo{benutzung durch braunschweig}
\todo{benutzung durch amsterdam}
